---
title: "Sampling bias in radiocarbon dating project: summarising the AustArch database"
author: "Rebecca Wheatley"
date: "8 September 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

Load required packages:
```{r, message = FALSE}
library(ggformula)
library(plyr)
library(tidyverse)
library(data.table)
library(rcarbon)
library(parallel)
```

Load the AustArch data and make sure relevant columns read in correctly:
```{r, message = FALSE}
AustArch <- read.csv("C:/Users/Bec/Work/Projects/Radiocarbon dating/GitHub/Data/Austarch_1-3_and_IDASQ_28Nov13-1.csv")
AustArch$AGE_NORM     <- as.numeric(AustArch$AGE_NORM)
AustArch$ERROR        <- as.numeric(AustArch$ERROR)
AustArch$LATITUDE     <- as.numeric(AustArch$LATITUDE)
AustArch$LONGITUDE    <- as.numeric(AustArch$LONGITUDE)
AustArch$IBRA_REGION  <- as.factor(AustArch$IBRA_REGION)
AustArch$SITE         <- as.factor(AustArch$SITE)
```

Get rid of:  
* "unusable" samples 
* samples missing an age and/or error estimate  
* samples generated using any method that is not radiocarbon dating 
* samples that fall outside of the calibration age range (this should be anything < 400 years BP for marine samples, and < 200 years for terrestrial samples) 
* samples that fall outside of the age range we are interested in (in this case, we only want Holocene dates)
```{r, message = FALSE, results = 'hide'}
data <- AustArch %>%
  dplyr::filter(Data.pertinent.for.time.series.analysis.or.calibration != "Unusuable") %>%
  dplyr::filter(AGE_NORM != "NA") %>%
  dplyr::filter(ERROR != "NA") %>%
  dplyr::filter(METHOD == "Radiocarbon") %>%
  dplyr::filter((AGE_NORM >= 400 & Data.pertinent.for.time.series.analysis.or.calibration == "Marine") 
                | (AGE_NORM >= 200 & Data.pertinent.for.time.series.analysis.or.calibration == "Terrestrial")) %>%
  subset(AGE_NORM <= 12000)
```

Create a new column that records the number of dated samples available for that site: NO_DATES_PER_SITE
```{r, message = FALSE}
data <- data %>%
  dplyr::arrange(SITE) %>%
  dplyr::group_by(SITE, .drop = TRUE) %>%
  dplyr::mutate(DATES_PER_SITE = length(AGE_NORM))
glimpse(data)
```

There are 1,529 sites in our Holocene data set.

How many sites contain only one sample?
```{r, message = FALSE}
data.1sample = data %>% subset(DATES_PER_SITE == 1)
glimpse(data.1sample)
```
There are 823 sites that contain only 1 sample (~54% of the total sites)

How many sites contain 2 samples?
```{r, message = FALSE}
data.2samples <- data %>% 
  subset(DATES_PER_SITE == 2)
glimpse(data.2samples)
```
299 sites contain only 2 samples (~20% of sites)

How many sites contain 3 samples?
```{r, message = FALSE}
data.3samples <- data %>% 
  subset(DATES_PER_SITE == 3)
glimpse(data.3samples)
```
129 sites contain only 3 samples (~8% of sites)

How many sites contain 5 or more samples?
```{r, message = FALSE}
data.5samples <- data %>% 
  subset(DATES_PER_SITE >= 5)
glimpse(data.5samples)
```
181 sites contain 5 or more samples (~12% of sites)